---
title: "Advanced R Topics"
description: |
site: distill::distill_website
output: 
    distill::distill_article:
       toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```

## Overview

Welcome to FES/FOR 505, Advanced R Topics. This is an informal class for students wanting to get a little extra practice in R.  We will work together through material in real time.

The class will meet once per week for 2 hours over 10 weeks. Each class will be a mix of instruction on the topic, discussion, and active work in R.  Materials (slides, links, R scripts) will be listed on the [Materials](materials.html) page throughout the quarter, so check back each week.

The plan so far:   
Week 01: Introduction to Git/GitHub  
Week 02: Build a personal website with **distill** and **postcards**  
Week 03: Deploy website using GitHub Pages  
Week 04: Explore **gt** for making tables  
Week 05: **gt** flair: colors, images, and themes    
Week 06: Learn to write functions in R  
Week 07: More Git: collaborators, merge conflicts, and pull requests  
Week 08: Interactive graphics and tables     
Week 09: Reproducible examples with **reprex**      
Week 10: GitHub issues    

## Software

We will be working in R via RStudio. Make sure you have up-to-date versions of both of these. Current versions:  
**R 4.0.4**  
**RStudio 1.4**  

## Packages

I'll be adding to the package list throughout the term as we hammer out exactly what topics will be covered. You'll want to install these on your computer so you're ready to go. I've added the weeks each package is first used in parentheses.

**distill v. 1.2** (*Week 2*)  
**postcards v. 0.2.0** (*Week 2*)  
**devtools v. 2.3.2** (*Week 2*)  
**fontawesome v. 0.1.0** (not 0.2.0; *Week 2*)  
**gt v. 0.2.2**  (*Week 4*)  
**webshot v. 0.5.2** (*Week 4*)  
**dplyr v. 1.0.5 ** (*Week 4*)  
**tidyr v. 1.1.3** (*Week 5*)  
**usethis v. 2.0.1** (*Week 7*)  
**gh v. 1.2.1** (*Week 7*)  
**htmlwidgets v. 1.5.3** (*Week 8*)  
**ggplot2 v. 3.3.3** (*Week 8*)  
**palmerpenguins v. 0.1.0** (*Week 8*)  
**plotly v. 4.9.3** (*Week 8*)   
**DT v. 0.18** (*Week 8*)  
**leaflet v. 2.0.4.1** (*Week 8*)  
**dygraphs v. 1.1.1.6** (*Week 8*)  
**flexdashboard v. 0.5.2** (*Week 8*)  
**reprex v. 2.0.0** (*Week 9*)  
**datapasta v. 3.1.0** (*Week 9*)  


Code for installing all packages:  

```{r, eval = FALSE, echo = TRUE}
install.packages(c("distill", "postcards", "devtools", "fontawesome", 
                   "gt", "webshot", "dplyr", "tidyr", "usethis",
                   "gh", "htmlwidgets", "ggplot2", "palmerpenguins",
                   "plotly", "DT", "leaflet", "dygraphs", "flexdashboard",
                   "reprex", "datapasta"))
```

