---
title: "Interactive elements in HTML documents"
output: 
  html_document:
      toc: TRUE
      toc_float: TRUE
      df_print: "paged"
date: "May 18, 2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use this document to add code and knit to explore options for adding interactive elements to HTML documents.  

Unlike in previous weeks, not all code is present in the document yet. You will be copy-pasting or typing in code to some of the code chunks from the slides. Then knit and explore HTML output document.

You can find today's slides [here](https://aosmith16.github.io/spring-r-topics/slides/week08_interactive_tables.html#1). The slides contain introductory material as well as a list of resources on the various **htmlwidgets** we are going to explore that aren't included here.

## R packages

Be sure to run code chunks in order, starting with loading packages.

Load **palmerpenguins 0.2.2** and **ggplot2 1.0.5** now.

We'll be loading additional packages as we go so we are clear on which package we are using in each section.  

```{r, message = FALSE, warning = FALSE}
library(palmerpenguins) # v. 0.1.0
library(ggplot2) # v. 3.3.3
```

```{r cars}
library(plotly)
library(ggplot2)
library(palmerpenguins)
g1 = ggplot(data = penguins, aes(x = bill_depth_mm, 
                                 y = bill_length_mm, 
                                 color = species)) +
    geom_point()
ggplotly(g1)
```

